---
title: "Visualization"
subtitle: "Part II"
output: github_document
---

```{r setup}
library(tidyverse)
library(patchwork)
```

## Load weather data

```{r load weather}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728", "USC00519397", "USS0023B17S"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(
      id, 
      USW00094728 = "CentralPark_NY", 
      USC00519397 = "Waikiki_HA",
      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())

weather_df
```

## Labels

```{r labelling}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = 0.5) +
  labs(
    title = "Temperature plot",
    x = "Minimum daily temp",
    y = "Maximum daily temp",
    caption = "Data from rnoaa package, temperatures from 2017"
  )
```

## Scales

```{r scales}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = 0.5) +
  labs(
    title = "Temperature plot",
    x = "Minimum daily temp",
    y = "Maximum daily temp",
    caption = "Data from rnoaa package, temperatures from 2017"
  ) +
  scale_x_continuous(
    breaks = c(-15,0,15)
  ) +
  scale_y_continuous(
    trans = "sqrt",
    position = "right"
  )
```
Can use trans = to transform, can do sqrt, log, and can use position to choose location of Y-axis

#### Introducing scale_color_hue

```{r hue}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = 0.5) +
  labs(
    title = "Temperature plot",
    x = "Minimum daily temp",
    y = "Maximum daily temp",
    caption = "Data from rnoaa package, temperatures from 2017"
  ) +
  scale_color_hue(
    name = "Location",
    h = c(100,300))
```
name = in this scale_color_hue function is the name of the legend, so we rename to "Location".

#### Introducing scale_color_viridis

```{r viridis}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = 0.5) +
  labs(
    title = "Temperature plot",
    x = "Minimum daily temp",
    y = "Maximum daily temp",
    caption = "Data from rnoaa package, temperatures from 2017"
  ) +
  viridis::scale_color_viridis(
    name = "Location",
    discrete = TRUE
  )
```
Viridis is better for those who are color blind, easier than creating our own color scales (which Jeff does not recommend doing)





