---
title: "Visualization"
subtitle: "Part II"
output: github_document
---

```{r setup}
library(tidyverse)
library(patchwork)
```

## Load weather data

```{r load weather}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728", "USC00519397", "USS0023B17S"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(
      id, 
      USW00094728 = "CentralPark_NY", 
      USC00519397 = "Waikiki_HA",
      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())

weather_df
```

## Labels

```{r labelling}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = 0.5) +
  labs(
    title = "Temperature plot",
    x = "Minimum daily temp",
    y = "Maximum daily temp",
    caption = "Data from rnoaa package, temperatures from 2017"
  )
```

## Scales

```{r scales}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = 0.5) +
  labs(
    title = "Temperature plot",
    x = "Minimum daily temp",
    y = "Maximum daily temp",
    caption = "Data from rnoaa package, temperatures from 2017"
  ) +
  scale_x_continuous(
    breaks = c(-15,0,15)
  ) +
  scale_y_continuous(
    trans = "sqrt",
    position = "right"
  )
```
Can use trans = to transform, can do sqrt, log, and can use position to choose location of Y-axis

#### Introducing scale_color_hue

```{r hue}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = 0.5) +
  labs(
    title = "Temperature plot",
    x = "Minimum daily temp",
    y = "Maximum daily temp",
    caption = "Data from rnoaa package, temperatures from 2017"
  ) +
  scale_color_hue(
    name = "Location",
    h = c(100,300))
```
name = in this scale_color_hue function is the name of the legend, so we rename to "Location".

#### Introducing scale_color_viridis

```{r viridis}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = 0.5) +
  labs(
    title = "Temperature plot",
    x = "Minimum daily temp",
    y = "Maximum daily temp",
    caption = "Data from rnoaa package, temperatures from 2017"
  ) +
  viridis::scale_color_viridis(
    name = "Location",
    discrete = TRUE
  )
```
Viridis is better for those who are color blind, easier than creating our own color scales (which Jeff does not recommend doing)

## Themes

#### Pre-defined themes exist and are useful/nice looking

```{r themes}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = 0.5) +
  labs(
    title = "Temperature plot",
    x = "Minimum daily temp",
    y = "Maximum daily temp",
    caption = "Data from rnoaa package, temperatures from 2017"
  ) +
  viridis::scale_color_viridis(
    name = "Location",
    discrete = TRUE
  )+
  theme_minimal()
```
Can use other themes, but this removes the default aspects of the theme and use theme_minimal.

#### Can use the function theme() to do some of your own adjustments.

```{r move legend}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = 0.5) +
  labs(
    title = "Temperature plot",
    x = "Minimum daily temp",
    y = "Maximum daily temp",
    caption = "Data from rnoaa package, temperatures from 2017"
  ) +
  viridis::scale_color_viridis(
    name = "Location",
    discrete = TRUE
  ) +
  theme(legend.position = "bottom")
```

### Important to note: if you put a theme (example theme_minimal) after you put another theme function (like changing the legend location), it will go to the default for the theme_minimal. You need to include default theme types before you include the theme elements.

## Setting options that Jeff likes

```{r, eval = FALSE}
library(tidyverse)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```
Will include at the top of all R markdown documents to keep formatting the same for any plots created in any fuutre R markdown document.

## Data arguments in `geom`

```{r data arguments}
central_park = 
  weather_df %>% 
  filter(name == "CentralPark_NY")

waikiki = 
  weather_df %>% 
  filter(name == "Waikiki_HA")

ggplot(data = waikiki, aes(x = date, y = tmax, color = name)) +
  geom_point() +
  geom_line(data = central_park)
```

## Time for patchwork

```{r patchwork}
tmin_tmax_p = 
  weather_df %>% 
    ggplot(aes(x = tmin, y = tmax, color = name)) +
    geom_point(alpha = 0.5) +
    theme(legend.position = "bottom")

prcp_dens_p = 
  weather_df %>% 
  filter(prcp > 0) %>% 
  ggplot(aes(x = prcp, fill = name)) +
  geom_density(alpha = 0.5)

tmax_date_p = 
  weather_df %>% 
  ggplot(aes(x = date, y = tmax, color = name)) +
  geom_point() +
  geom_smooth() +
  theme(legend.position = "none")

tmin_tmax_p + prcp_dens_p
```
Patchwork displays when you name these plots and then use + or / to determine positioning and inclusion of plots.

## Data manipulation

```{r manipulate}
weather_df %>% 
  mutate(
    name = factor(name),
    name = forcats::fct_relevel(name, c("Waikiki_HA"))
  ) %>% 
  ggplot(aes(x = name, y = tmax, fill = name)) +
  geom_violin(alpha = 0.5)
```

#### Getting densities for tmin and tmax overlaid.

```{r tmin and tmax}
weather_df %>% 
  pivot_longer(
    tmax:tmin,
    names_to = "observation",
    values_to = "temperatures"
  ) %>% 
  ggplot(aes(x = temperatures, fill = observation)) +
  geom_density(alpha = 0.5) +
  facet_grid(. ~ name)
```


