---
title: "Visualization"
subtitle: "Part I"
output: github_document
---

```{r setup}
library(tidyverse)
library(ggridges)
```

## Load weather data

```{r load weather data}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728", "USC00519397", "USS0023B17S"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(
      id, 
      USW00094728 = "CentralPark_NY", 
      USC00519397 = "Waikiki_HA",
      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())
```

## Scatterplots

#### First scatterplot!

```{r scatterplot1}
ggplot(weather_df, aes(x = tmin, y = tmax)) +
  geom_point()
```

#### New approach, same plot

```{r scatterplot2}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) +
  geom_point()
```
Allows for some data manipulation before graphing in the same step, creates consistency between this and how we code other things in R.

#### Save and edit a plot object

```{r plot object}
weather_plot = 
  weather_df %>% 
  ggplot(aes(x = tmin, y = tmax))

weather_plot + geom_point()
```
Allows you to make different types of graphs on the same plot.

## Advanced scatterplot

Let's make it fancy.

```{r fancy scatterplot}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point() +
  geom_smooth(se = FALSE)
```
Can also place aes in the other geometries, but if you place it in the ggplot function, then you will apply that aesthetic to everything that comes later.

#### Faceting some things.

```{r faceting}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = 0.2, size = 0.8) +
  geom_smooth(se = FALSE) +
  facet_grid(. ~ name)
```
The facet grid function says (rows ~ columns), so in this case, we are saying no rows, just columns, defined by the variable name.
Alpha indicates the level of transparency, size sets the size of the points. Alpha can also be used in ggplot and can be set by variables (example: alpha = tmin).

#### Time to combine some elements and create a new plot.

```{r another plot}
weather_df %>% 
  ggplot(aes(x = date, y = tmax, color = name)) +
  geom_point(aes(size = prcp), alpha = 0.5) +
  geom_smooth(se = FALSE) +
  facet_grid(. ~ name)
```

## A little note

You can have whatever geoms you want.

There are some fun ones `geom_density2d()`, `geom_hex`.

## Univariate plots.

Start off with histograms.

```{r histogram}
weather_df %>% 
  ggplot(aes(x = tmin)) +
  geom_histogram()
```

Some things to do with histograms.

```{r histogram2}
weather_df %>% 
  ggplot(aes(x = tmin, fill = name)) +
  geom_histogram(position = "dodge")
```
We use fill= rather than color= to fill in the bars of the histogram. Using position = "dodge" means that the bars will not stack on top of each other.
Could also use facet_grid() to see the three plots separately.

#### New geometry - geom_density()

```{r geom_density}
weather_df %>% 
  ggplot(aes(x = tmin, fill = name)) +
  geom_density(alpha = 0.3)
```

#### Box plots

```{r boxplot baby}
weather_df %>% 
  ggplot(aes(x = name, y = tmin)) +
  geom_boxplot()
```

#### Trendy plots

```{r violin}
weather_df %>% 
  ggplot(aes(x = name, y = tmin, fill = name)) +
  geom_violin(alpha = 0.5) +
  stat_summary(fun = "median")
```

```{r ridge plot}
weather_df %>% 
  ggplot(aes(x = tmin, y = name)) +
  geom_density_ridges()
```
Need to load the ggridges packages in order to use this plot.

## Save and embed

#### Let's save a scatterplot.

```{r}
weather_plot = 
  weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point()

ggsave("weather_plot.pdf", weather_plot, width = 8, height = 5)
```

#### Let's embed

```{r, fig.width=12, fig.height=6}
weather_plot
```

